---
title: "Litter Size Evolution pre-run"
output: html_document
---

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ape)
library(ggtree)
library(knitr)
library(phytools)
```
## Plot the consensus tree

```{r fig.width=15, fig.height=60 } 
  tree_readin = read.nexus('1028_output.nex')
  tree2 = random.tree<-sample(tree_readin,size=1)[[1]]
  
  print(tree2)
  plot(tree2)

  consensusTree = consensus(tree_readin, p = 0.5, check.labels = FALSE)
  plot(consensusTree)

 # kable(charstate)
```

## Ancestral State Reconstruction
I was able to do a basic reconstruction using phytools.I can see the projected litter size for ancestral nodes. 
Next steps:

1. Do the analysis on a consensus tree instead of a random tree

2. Clearly know projected nuemerical value for ancestral nodes 

3. Figure out a better plotting scheme, the rainbow colorscale is not visually informative. 

````{r fig.width=15, fig.height=60 } 
# Character state matrix 
charstate = read.csv('1028_MVP2.csv', row.names = 1)
charstate = as.matrix(charstate)[,1]

plotTree(tree2, type = 'fan', ftype = 'i', fsize = 1)
result = ace(charstate, tree2, type = 'continuous', method = 'ML')
class(charstate)

```

````{r fig.width=15, fig.height= 80 } 

contresult = contMap(tree2, charstate, plot = FALSE)
plot(contresult, lwd = 5, legend = 0.7 * max(nodeHeights(tree2)),outline = 0) # fsize = c(0.5, 0.5)

# D = data.frame(
#     species = c(charstate
# )
# 
# mass_plot = ggtree( tree2 ) %<+% D + 
#   geom_tiplab( fontface = "italic", offset=20 ) +
#   xlim(0, 300) +
#   geom_tippoint( aes(color= x), size=3, alpha=1 )
# 
# mass_plot


```
