---
title: "Litter Size Evolution"
output:
  pdf_document: default
  html_document: default
---

Consensus tree extracted from an Upham et al. paper (Jetz lab)  
Data: Out of 531 data points, 458 are in the consensus tree (thus plotted), out of 1200 species searched (2200 in total)  

Acomys cahirinus (menstruating cairo spiny mouse) is included, it might take a few seconds to find.

For the rest 73 data points (531-458), I don't know where they lie on the tree  
next steps:  
1. find out where the 73 data points lie, and modify the consensus tree to include them   
2. keep searching species   
3. get more familiar with rodentia phylgoeny  


```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ape)
library(ggtree)
library(knitr)
library(phytools)
library(tidyverse)
library(magrittr)
```

```{r fig.width=15, fig.height = 200, include = FALSE} 
  tree_global = read.tree('RAxML_bipartitions.result_FIN4_raw_rooted_wBoots_4098mam1out_OK.newick')
  n_tips = length(tree_global$tip.label)
  root_node = tree_global$edge[(!tree_global$edge[,1] %in% tree_global$edge[,2]),1] %>% unique()
  root_age = max( dist.nodes(tree_global)[,root_node] )
  
  label_offset = 0.02 * root_age
  tree_width = 5 * root_age
```

```{r, include = FALSE}
## drop all the non-relevant tips

charstate = read.csv('1111_FP_upTo1200.csv') %>%
  mutate(Average = as.numeric(Average))

tips <- tree_global$tip.label
tips_short = str_replace(tips, "[A-Z_]+$", "")
tree_global$tip.label = tips_short

tip_names = data.frame(Name_long=tips, Name=tips_short)
charstate %<>% left_join(tip_names)

missing_in_charstate = tree_global$tip.label[! tree_global$tip.label %in% charstate$Name]
 
tree_global = ape::drop.tip(tree_global, missing_in_charstate)

charstate %<>% filter(!is.na(Name_long))

rank <- match( charstate[,1], tree_global$tip.label )

charstate = charstate[match( tree_global$tip.label, charstate[,1] ), ]

 if(! all( charstate[,1] ==tree_global$tip.label ) ){
      stop("Error: tip labels and dataframe rows not in same order!")
    }

```

```{r,  fig.height = 200, fig.width = 40, echo = FALSE}


litter_ancestral = 
  ace(
  charstate$Average, 
  tree_global,
  type = "continuous",
  method = "pic",
  model = "BM"
)

# geo_ggtree = ggtree( geoTree ) + 
#   geom_tiplab( fontface = "italic", offset=label_offset ) +
#   xlim(0, tree_width) +
#   geom_text2(aes(label=node), col="red", nudge_x=label_offset/2 ) +
#   geom_tippoint( aes(color=node_values), size=3, alpha=1 )
# geo_ggtree

# interval_width = wing_ancestral$CI95[,2] - wing_ancestral$CI95[,1]
# node_values = c( rep(NA, nrow(geoData)), interval_width )
# kable( data.frame( node = seq(1, length(node_values)), width=node_values ) )
# geo_ggtree + 
#   geom_nodepoint(  aes(color=node_values), size=3, alpha=1 )

node_values = c(charstate$Average, round(litter_ancestral$ace, digits= 2))

globaltree = ggtree(tree_global ) + 
  geom_tiplab( fontface = "italic", size= 10, offset = label_offset) + 
  xlim(0, 1.2) + 
  geom_text2(aes(label = node_values), nudge_x = 0.02, size = 8) +
  geom_tippoint(aes(color=node_values), size= 8, alpha=1 ) + 
  scale_color_gradient(low = "red", high = "green") +
  theme(legend.text=element_text(size= 40),
        legend.title=element_text(size=60),
        legend.key.size= unit(3,"cm"), 
        legend.position= "top")

globaltree

```





