---
title: "Litter Size Evolution"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

Consensus tree extracted from an Upham et al. paper (Jetz lab)  
Data: 1006 out of 2278 species


```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ape)
library(ggtree)
library(knitr)
library(phytools)
library(tidyverse)
library(magrittr)
```

```{r fig.width=15, fig.height = 200, include = FALSE} 
  tree_global = read.tree('RAxML_bipartitions.result_FIN4_raw_rooted_wBoots_4098mam1out_OK.newick')
  n_tips = length(tree_global$tip.label)
  root_node = tree_global$edge[(!tree_global$edge[,1] %in% tree_global$edge[,2]),1] %>% unique()
  root_age = max( dist.nodes(tree_global)[,root_node] )
  
  label_offset = 0.02 * root_age
  tree_width = 5 * root_age
```

```{r, include = FALSE}
## drop all the non-relevant tips

charstate = read.csv('1128_subset.csv') 
charstate
charstate = charstate%>%mutate(Average = as.numeric(Average))

tips <- tree_global$tip.label
tips_short = str_replace(tips, "[A-Z_]+$", "")
tree_global$tip.label = tips_short

tip_names = data.frame(Name_long=tips, Name=tips_short)
charstate %<>% left_join(tip_names)

missing_in_charstate = tree_global$tip.label[! tree_global$tip.label %in% charstate$Name]
 
tree_global = ape::drop.tip(tree_global, missing_in_charstate)

charstate %<>% filter(!is.na(Name_long))

rank <- match( charstate[,1], tree_global$tip.label )

charstate = charstate[match( tree_global$tip.label, charstate[,1] ), ] #the data order matches, checked manually as well.

 if(! all( charstate[,1] ==tree_global$tip.label ) ){
      stop("Error: tip labels and dataframe rows not in same order!")
    }

```

```{r,  fig.height = 200, fig.width = 40, echo = FALSE}


litter_ancestral = 
  ace(
  charstate$Average, 
  tree_global,
  type = "continuous",
  method = "REML",
  model = "BM"
)

node_values = c(charstate$Average, round(litter_ancestral$ace, digits= 2))
node_values

globaltree = ggtree(tree_global, layout = 'circular' ) + 
  geom_tiplab( fontface = "italic", size= 5, offset = label_offset) + 
  xlim(0, 1.2) + 
  geom_text2(aes(label = node_values), nudge_x = 0.02, size = 4) +
  geom_tippoint(aes(color=node_values), size= 4, alpha=1 ) + 
  geom_nodepoint( aes(color=node_values), size= 4, alpha=1 ) + # problem here! I have trouble adding internal node points. 
  scale_color_gradient(low = "red", high = "green") +
  theme(legend.text=element_text(size= 40),
        legend.title=element_text(size=60),
        legend.key.size= unit(3,"cm"), 
        legend.position= "top")

globaltree

##### The following are for pic method, not sure to include or not 

# litter_ancestral_pic = 
#   ace(
#   charstate$Average, 
#   tree_global,
#   type = "continuous",
#   method = "pic",
#   model = "BM"
# )

# node_values_pic = c(charstate$Average, round(litter_ancestral_pic$ace, digits= 2))
# 
# globaltree_pic = ggtree(tree_global ) + 
#   geom_tiplab( fontface = "italic", size= 10, offset = label_offset) + 
#   xlim(0, 1.2) + 
#   geom_text2(aes(label = node_values_pic), nudge_x = 0.02, size = 8) +
#   geom_tippoint(aes(color=node_values_pic), size= 8, alpha=1 ) + 
#   scale_color_gradient(low = "red", high = "green") +
#   theme(legend.text=element_text(size= 40),
#         legend.title=element_text(size=60),
#         legend.key.size= unit(3,"cm"), 
#         legend.position= "top")
# 
# globaltree_pic


```

This gives us a variety of information about the ancestral character reconstructions:

- `$ace` contains the reconstructed ancestral character states. There is one value per internal node, and they are in the order that the internal nodes are numbered in.
- `$sigma2` is the estimate of the Brownian Motion parameter.
- `$CI95` are the confidence intervales on the ancestral character state reconstructions.

